---
title: "DATA607 - Assignment 2:"
author: "Kailot C. Harris"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Assignment Details

Assignment: SQL and R – Movie Ratings
Objective
Collect simple movie-rating data, store it in a SQL database, and analyze it in R.

Implementation Notes:
PostgreSQL is the recommended (and supported) relational database for this assignment.  If you are unable to install and configure PostgreSQL in your environment, you may find SQLite easier to use.  Students experienced with relational databases may use any local or cloud-based database for their assignment.
NEVER include passwords in your code.  For this assignment, the ability to reproduce your work in my environment is not required, but you need to provide all of the necessary code, include the code used to create and populate tables.
Task Description

Select six recent popular movies (or television episodes or books or songs or ...)

Ask at least five people to rate each movie they have seen on a 1–5 scale.

Store the collected ratings in a SQL database of your choice.

Load the data from SQL into R as a dataframe.

Implementation Notes

The assignment is intentionally open-ended; multiple approaches are acceptable.

You may:

Query the SQL database directly from R, or

Export SQL data to a .csv file and load it into R.

If you use a GUI to create or populate tables, you must also provide the SQL code needed to recreate them.

Reproducibility

Your solution should be reproducible assuming the same SQL server and R environment.

You may remove or mask database passwords.

Full reproducibility of survey responses is not required.

Missing Data

You must demonstrate a reasonable strategy for handling missing ratings, since not all participants will have seen all movies.

Optional Enhancements
You may improve your solution by:

Using survey software to collect ratings

Managing database credentials securely

Designing a normalized database schema

Exploring whether standardizing ratings is beneficial

Submission Requirements

Post all code (SQL, R, Quarto) in a GitHub repository

Submit the GitHub link

RPubs publication is not required

Group work is allowed, but each student must submit individually and list collaborators

Professional Expectation
Treat this as portfolio-quality work suitable for a job interview.

2. Optional Instructions to Make This Easier for Students New to SQL and Recommender Systems
A. Suggested Conceptual Roadmap
Consider approaching the assignment in three layers:

Data Collection (Problem Framing)

Think of this as a mini user-item ratings dataset.

Users = people you survey

Items = movies

Ratings = integers from 1–5 (with missing values allowed)

Data Storage (SQL Fundamentals)

Start with two or three tables, not one:

users(user_id, name)

movies(movie_id, title)

ratings(user_id, movie_id, rating)

Learn:

Primary keys

Foreign keys

Why many-to-many relationships require a junction table

Data Analysis (R Fundamentals)

Use R primarily for:

Data inspection

Cleaning missing values

Simple summaries (means, counts)

At this stage, you are not building a full recommender—only preparing the data.

B. Concrete Starter Checklist
SQL

Create tables using CREATE TABLE

Insert data using INSERT INTO

Use SELECT with JOIN to combine tables

Verify missing ratings using NULL

R

Use DBI / RSQLite or RPostgres or read a CSV

Convert SQL output into a tidy dataframe

Use dplyr to:

Count ratings per user and movie

Compute average ratings

Recommender Systems (Conceptual Only)

Understand that this dataset resembles:

A user–item matrix

The raw input for collaborative filtering

No advanced algorithms are required yet

## Approach

Type out approach here

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure}

```